---
title: "Irish Times"
output: html_document
---

```{r setup, include=FALSE}
                      
```

## R Introduction 

For this dataset I looked at 1.42 million news headlines in the Irish Times published from 1996 to the end of 2018: In this data set I will attempt to provided fundamental analysis on the data and make visual plots on any topics that I find interesting. 

```{r packages, include = FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(forcats)
library(data.table)
library(ggplot2)
library(stringr)
```

##data
data can be found [here](https://www.kaggle.com/therohk/ireland-historical-news). I renamed irishtimes-date-text into irish1 and the other into irish2. 

```{r data, include =FALSE}
irish1 <- read.csv("irish1.csv")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r summary, include = FALSE}
summary(irish1)
searchterm <- c("business")
irish1 %>% count(headline_category, sort = TRUE) -> sort_by_category
head(sort_by_category)
tail(sort_by_category)
#testing if the code work
#using on the full categories
list(sub("\\..*", "", irish1$headline_category)) -> sums1
data.frame(table(sums1))-> headline_general
head(headline_general)
# a more elegant approach
mutate(irish1, cat = sub("\\..*", "", irish1$headline_category)) -> irish1edit
#dont need this code atm
#class(headline_general)

#making a rough graph
headline_general %>% ggplot(aes(x = sums1, y = Freq)) + geom_bar(stat="Identity", fill = "steelblue") + labs(x = "Cateogry", y = "Frequency")  -> graph1
graph1

#proper graph using ggplot
irish1edit$publish_date <- as.Date(as.character(irish1edit$publish_date), format = "%Y%m%d")
irish1edit %>% group_by(publish_date, cat) %>% count() %>% ggplot(aes(x= publish_date, y = n, color = cat)) + geom_line()+ geom_point() + scale_x_date(limits = as.Date(c('1996-01-02', '1997-02-01')))
class(irish1edit$publish_date)

##attempt 2 looking only at year and month 

irish1edit1 <- irish1edit
irish1edit1$publish_date <-format(irish1edit1$publish_date, format = "%Y-%m")
head(irish1edit1)
irish1edit1 %>% group_by(publish_date, cat) %>% count() -> irish_data_cleaned
head(dput(irish_data_cleaned))

#attempts to clean the data and brush up the limits. 
irish_data_cleaned$publish_date1 <- as.Date(irish_data_cleaned$publish_date, format = "%Y-%m")
irish_data_cleaned %>% ggplot(aes(x = publish_date, y = n, group = cat))  + geom_line(aes(color = cat)) + scale_x_discrete(name = "Publish Date", breaks = 10)
#+ scale_x_date(date_breaks = "1 month", date_labels = "%Y-%m")
class(irish_data_cleaned$publish_date)
#+ xlim('1996-01', '1999-02')
```
##Results
Here we can see the overal summary of the data. By printing the first and last six rows of the data we can see that there is a high disrepancy between the most popular categories and the least popular. We can also see that the categories are further divided by subcategories, where some subcategories are more popular the main categories. Here we can see that opinion.letters are more popular than opinion. We can see that news is the most popular category for the newspaper followed by sport and then closely by business. 