---
title: "Irish Times"
output: html_document
---

```{r setup, include=FALSE}
                      
```

## Introduction 

For this dataset I looked at 1.42 million news headlines in the Irish Times published from 1996 to the end of 2018: In this data set I will attempt to provided fundamental analysis on the data and make visual plots on performance of categories over time in the print run of this newspaper. 

```{r packages, include = FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(forcats)
library(data.table)
library(ggplot2)
library(stringr)
```

##Data
The data can be found [here](https://www.kaggle.com/therohk/ireland-historical-news).

```{r data, include =FALSE}
irish1 <- read.csv("irish1.csv")
```


```{r summary, include = FALSE}
summary(irish1)
searchterm <- c("business")
irish1 %>% count(headline_category, sort = TRUE) -> sort_by_category
#transforming into a Date class
irish1 <- transform(irish1, publish_date1 = as.Date(as.character(publish_date), "%Y%m%d"))
head(sort_by_category)
tail(sort_by_category)
#testing if the code work
#using on the full categories
list(sub("\\..*", "", irish1$headline_category)) -> sums1
data.frame(table(sums1))-> headline_general
head(headline_general)
# a more elegant approach
mutate(irish1, cat = sub("\\..*", "", irish1$headline_category)) -> irish1edit
#dont need this code atm
#class(headline_general)

#proper graph using ggplot

irish1edit %>% group_by(publish_date1, cat) %>% count() -> counted_irish
counted_irish %>% ggplot(aes(x= publish_date1, y = n, color = cat)) + geom_line() + scale_x_date(limits = as.Date(c('1996-01-01', '1996-12-31'))) + xlab ("Date") + ylab("Counts of Published Articles") + ggtitle("Category Performance in the First Year") -> graph_first_year

counted_irish %>% ggplot(aes(x= publish_date1, y = n, color = cat)) + geom_line(alpha = 1/2) +  xlab("Date") + ylab("Counts of Published Articles") + ggtitle("Overall View of Category Performance in Irish Time's Entire Print Run") -> graph_overall

sort_by_category %>% ggplot(aes(x = reorder(headline_category, -n), y = n)) + geom_col() + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + xlab("Categories") + ylab("Counts of Categories") + ggtitle("Graph of Categories from Most Popular to Least") -> graph_categories

##attempt 2 looking only at year and month 

# irish1edit1 <- irish1edit
# 
# irish1edit1$publish_date <-format(irish1edit1$publish_date, format = "%Y-%m")
# 
# head(irish1edit1)
# 
# irish1edit1 %>% group_by(publish_date, cat) %>% count() -> irish_data_cleaned
# 
# head(dput(irish_data_cleaned))
# 
# #attempts to clean the data and brush up the limits. 
# irish_data_cleaned$publish_date1 <- as.Date(irish_data_cleaned$publish_date, format = "%Y-%m")
# irish_data_cleaned %>% ggplot(aes(x = publish_date, y = n, group = cat))  + geom_line(aes(color = cat)) 
# 
# #+ scale_x_discrete(name = "Publish Date", breaks = 10)
# #+ scale_x_date(date_breaks = "1 month", date_labels = "%Y-%m")
# class(irish_data_cleaned$publish_date)
# #+ xlim('1996-01', '1999-02')
```


##Results
Here we can see the overal summary of the data. By printing the first and last six rows of the data we can see that there is a high disrepancy between the most popular categories and the least popular. The most popular categories being news, sport, business and interestingly enough opinions.letters as a subcategory is more popular than the main category of opinions in general. Looking at the least popular catories we see that there are many that were printed a fewer of 10 times in the entire print run of irish times. In our plot we can see that the categories heavily drop off after the top five. I believe that by combining all sub categories in the main ones we can have a clear understanding of performance of categories over time. 
```{r plot, echo = TRUE}
head(sort_by_category)
tail(sort_by_category)
graph_categories


```

We can see that news is the most popular category for the newspaper followed by sport and then closely by business. With the graphical output it's easy to see that news is the most popular in the year 1996 of when data was collected and also coincidentally the first year of publication. I attempt to depict the popularity of news articles over the entire course of the Irish Time's publication run. While there are spikes in categories such as sports as perhaps to major sporting events or those events that are particularly popular within the readers of Irish Times, the news category remains to be the most popular. 
```{r plot2, echo = TRUE}
graph_first_year
graph_overall


```