---
title: "Mushroom"
output: html_document
---

```{r setup, include=FALSE}
```

##Introduction

Hello, this is my first real project that looks at [mushroom edibility data](https://www.kaggle.com/uciml/mushroom-classification/downloads/mushroom-classification.zip/1) from Kraggle. mushroom edibilty was deteremined from a multitude of factors provided in the data. Detailed descriptions of the data variables can be found on Kraggle. 

```{r load_packages, include= FALSE}
library(rpart)
library(rpart.plot)
library(tree)
library(randomForest)
library(caTools)
library(caret)
library(readr)
library(dplyr)
library(tidyr)
library(forcats)
library(data.table)
library(ggplot2)
```

## I installed the relevant packages needed, but the code has been omitted. 


```{r pressure, include = FALSE}
mushroom_data <- read.csv("mushrooms.csv")

```



```{r plot code, include = TRUE}
summary(mushroom_data)
placeholder <- cbind.data.frame(Var = names(mushroom_data), total = sapply(mushroom_data, function(x){as.numeric(length(levels(x)))}))
print(placeholder)
##veil type seems to be only 1 so we can drop
mushroom_data$veil.type <- NULL

##Cleaning Data
#splitting into training and test sets
n <- nrow(mushroom_data)
n_train <- round(0.8*n)
set.seed(123)
train_indices <- sample(1:n, n_train)
mushroom_train <- mushroom_data[train_indices,]
mushroom_test <- mushroom_data[-train_indices,]
x_train <- mushroom_train
x_test <- mushroom_test
x_test$class <- NULL
x_train$class <- NULL


# Create a stratified sample for repeated cv using multifolded
cv.10.folds<-createMultiFolds(mushroom_train$class,k=10,times=2)

# create a control object for repeated cv in caret
ctrl.1<-trainControl(method="repeatedcv",number=10,repeats=2,index=cv.10.folds)

rf.1.cv<-train(x=x_train,y=mushroom_train$class,method="rf",trControl=ctrl.1,tuneLength=3)

#variable importance using Random Forest

plot(varImp(rf.1.cv),main="Random Forest - Variable Importance Plot")

# shows 100% accuracy, lets predict on test set

y_predicted<-predict(rf.1.cv,x_test)

df1<-data.frame(Orig=mushroom_test$class,Pred=y_predicted)

confusionMatrix(table(df1$Orig,df1$Pred)) 

#__________________________________________________

cv.10.folds<-createMultiFolds(mushroom_train$class,k=10,times=2)
ctrl.1<-trainControl(method="repeatedcv",number=10,repeats=2,index=cv.10.folds)
mushroom_modelC <- train(x = x_train, y = mushroom_train$class, method = "rpart", trControl = ctrl.1, tuneLength = 5)
#looking at variable importannce using RPART
plot(varImp(mushroom_modelC),main="RPART - Variable Importance Plot")
rpart.plot(mushroom_modelC$finalModel)
```

##Results
The results from this random forest method and RPART showed us a 100% accuracy rating in predicting mushroom edibility from their factors. We can see the variable imporatance as shown by the graphs are identical in identifying that odor was the most important in determining the edibility of the mushroom followed by spore.print.color and gill. The decision tree is also given to show the factors as we move down the tree to determine edibility of mushrooms. The confusion matrix provided shows us the accuracy rating as mentioned before as all mushrooms that are edible are classified properly as edible and similarly for mushrooms that inedible. 





